---
title: "Hi! I am Sung"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

<link rel="stylesheet" href="https://cdnjs.cloudflare.com/ajax/libs/font-awesome/6.5.0/css/all.min.css">

<style>
.card-comp {
  background: #fdfdfd;
  padding: 22px 24px;
  border-radius: 8px;
  box-shadow: 0 2px 12px rgba(0,0,0,0.09);
  height: 100%;
}
.card-comp h3 {
  margin-top: 8px;
  font-size: 1.05em;
  color: #2c3e50;
}
.card-comp ul {
  padding-left: 16px;
  margin-bottom: 0;
  font-size: 0.92em;
  color: #444;
}
.card-comp .card-icon {
  font-size: 1.6em;
  color: #18bc9c;
  margin-bottom: 4px;
}
.card-comp-2 .card-icon { color: #2c7be5; }
.card-comp-3 .card-icon { color: #e67e22; }
.card-project { border: 1.5px solid #e8a0b0; }

.about-left {
  border-right: 1px solid #dee2e6;
  padding-right: 20px;
}

.pub-card {
  background: #fdfdfd;
  border: 1.5px solid #d0d5dd;
  border-radius: 8px;
  box-shadow: 0 2px 6px rgba(0,0,0,0.06);
  padding: 16px 18px;
  margin-bottom: 18px;
  height: 100%;
}
.pub-journal {
  font-size: 1.0em;
  font-weight: 700;
  color: #2c3e50;
  margin-bottom: 4px;
}
.pub-year {
  font-size: 0.85em;
  color: #18bc9c;
  font-weight: 600;
}
.pub-title {
  font-size: 0.80em;
  color: #666;
  line-height: 1.5;
  margin-top: 6px;
}
.pub-badge {
  font-size: 0.72em;
  color: #888;
  margin-top: 6px;
  font-style: italic;
}
.workflow-card {
  background: #fdfdfd;
  border: 1.5px solid #a8c4e0;
  border-radius: 8px;
  box-shadow: 0 2px 6px rgba(0,0,0,0.06);
  padding: 16px 18px;
  margin-bottom: 18px;
  height: 100%;
}
.about-right {
  flex: 1;
  min-width: 260px;
  background: #fdfdfd;
  padding: 22px 26px;
  border-radius: 8px;
  box-shadow: 0 2px 12px rgba(0,0,0,0.08);
}

.about-left h5 {
  font-size: 0.75em;
  text-transform: uppercase;
  letter-spacing: 0.08em;
  color: #999;
  margin: 16px 0 4px;
}
.about-left p {
  font-size: 0.85em;
  color: #555;
  line-height: 1.6;
  margin: 0;
}

h1 { font-size: 1.75em; }
h2 { font-size: 1.4em; }
h3 { font-size: 1.1em; }

</style>

---

# Sung Rye Park {#about}

<div class="about-right">

<span style="font-size:1.5em; font-weight:700;">COMPUTATIONAL BIOLOGIST</span>  
*Specializing in Translational Drug Discovery*

Computational Biologist with 7+ years delivering multi-omics integration and translational biomarker analyses in pharma collaborations (BMS, Takeda, Japan Tobacco)

Builds analytical frameworks that bridge biology, machine learning, and multi-omics data — designing systems-level approaches that make sense of heterogeneous biological datasets.
<br><br>

<div style="background:#f4f6f8; border-radius:6px; padding:12px 16px; font-size:0.88em; color:#555;">
**Domains:** Immuno-oncology · Metabolic disease (NAFLD) · Neurodegeneration (AD, TBI)  
**Expertise:** Single cell · Multi-omics integration · Machine Learning · Pipeline development · Analytical templates · Interactive reports
</div>

<br>
[CV](info/resume/SungryePark_resume.pdf){.btn .btn-primary} &nbsp;
[GitHub](https://github.com/sung2021){.btn .btn-default target="_blank"} &nbsp;
[LinkedIn](https://www.linkedin.com/in/sung-rye-park-8a130098/){.btn .btn-default target="_blank"}

</div>

---

# Featured Projects {#projects}

::: {.row}

::: {.col-md-6}
<div class="card-comp card-project" style="margin-bottom:20px;">
### **Novel function of Tle3: Multi-omics approach**
- Four modalities with pair-wise integration
- scRNA-seq, RNA-seq, ATAC-seq, Cut&Run (TLE3)
- *Nature Immunology* (2024); co-first author
<br><br>
`Pair-wise integration` `Immunology` `T cell` `Lineage`
</div>
:::

::: {.col-md-6}
<div class="card-comp card-project" style="margin-bottom:20px;">
### **STING Agonist Response: BMS Collaboration**
- Responding cell type stratification
- scRNA-seq + scTCR-seq
- *Cell Reports Medicine* (2024)
<br><br>
`Single-cell RNA-seq` `scTCR-seq` `Immunology` `Oncology` `Deconvolution`
</div>
:::

::: {.col-md-6}
<div class="card-comp card-project" style="margin-bottom:20px;">
### **Virtual Patient Cohort Generation (MicroIRAD)**

- CVAE framework learning TCGA-LUAD
- 400+ synthetic patients
- Patient-Model cell line pair similarity score (10,000+)
- Preclinical go/no-go decision support
<br><br>
`Deep Learning` `PyTorch/Tensorflow` `CVAE` `TCGA` `LUAD` `AutoEncoder`
</div>
:::

::: {.col-md-6}
<div class="card-comp card-project" style="margin-bottom:20px;">
### **Cell image + RNA-seq by MOFA2**

- Phenotypic profiling + Transcriptome
- Large scale compound response study
- End-to-end integration pipeline
<br><br>
`MOFA2` `Cell Painting` `Transcriptome` `Drug response` `Human lung cancer` `Multiomics`
</div>
:::

:::

---

# Selected Publications {#publications}

<div class="row">

<div class="col-md-4">
<div class="pub-card">
<div class="pub-journal">Nature Immunology <span class="pub-year">2024</span></div>
<div class="pub-title">The transcriptional cofactor Tle3 reciprocally controls effector and central memory CD8+ T cell fates.</div>
<div class="pub-badge">Co-first author &nbsp;·&nbsp; <a href="https://pmc.ncbi.nlm.nih.gov/articles/PMC10916363/" target="_blank">View Paper →</a></div>
</div>
</div>

<div class="col-md-4">
<div class="pub-card">
<div class="pub-journal">Cell Reports Medicine <span class="pub-year">2025</span></div>
<div class="pub-title">Immune targeting of triple-negative breast cancer through a clinically actionable STING agonist-CAR T cell platform.</div>
<div class="pub-badge"><a href="https://pmc.ncbi.nlm.nih.gov/articles/PMC12281382/" target="_blank">View Paper →</a></div>
</div>
</div>

<div class="col-md-4">
<div class="pub-card">
<div class="pub-journal">Cell <span class="pub-year">2021</span></div>
<div class="pub-title">Seq-Scope: Microscopic examination of spatial transcriptome.</div>
<div class="pub-badge"><a href="https://doi.org/10.1016/j.cell.2021.05.010" target="_blank">View Paper →</a></div>
</div>
</div>

</div>
<div class="row">
<div class="col-md-4">
<div class="pub-card">
<div class="pub-journal">Cell Reports <span class="pub-year">2021</span></div>
<div class="pub-title">Single cell transcriptome analysis of colon cancer cell response to 5-fluorouracil-induced DNA damage.</div>
<div class="pub-badge">First author &nbsp;·&nbsp; <a href="https://doi.org/10.1016/j.celrep.2020.108077" target="_blank">View Paper →</a></div>
</div>
</div>

<div class="col-md-4">
<div class="pub-card">
<div class="pub-journal">AJP-Endocrinology and Metabolism <span class="pub-year">2020</span></div>
<div class="pub-title">Holistic Characterization of Single Hepatocyte Transcriptome Responses to High Fat Diet.</div>
<div class="pub-badge">First author &nbsp;·&nbsp; <a href="https://pmc.ncbi.nlm.nih.gov/articles/PMC8260362/" target="_blank">View Paper →</a></div>
</div>
</div>

</div>

---

# Analysis Workflows {#workflows}

<div class="row">

<div class="col-md-4">
<div class="workflow-card">
<div class="pub-journal">Single Cell RNA-seq</div>
<div class="pub-title">QC, integration (Harmony), UMAP/clustering, cell type annotation, trajectory, AUC, scTCR-seq preprocessing</div>
<div class="pub-badge"><a href="https://sung2021.github.io/samplePages/scRNAseq.html" target="_blank">View Workflow →</a></div>
</div>
</div>

<div class="col-md-4">
<div class="workflow-card">
<div class="pub-journal">Bulk RNA-seq</div>
<div class="pub-title">PCA, TPM, DEG analysis, GSEA/ssGSEA, K-means clustering, feature selection</div>
<div class="pub-badge"><a href="https://sung2021.github.io/samplePages/RNAseq.html" target="_blank">View Workflow →</a></div>
</div>
</div>

<div class="col-md-4">
<div class="workflow-card">
<div class="pub-journal">NK cell identification using ML</div>
<div class="pub-title">Elastic Net for feature selection, K-means + DEG-based pattern identification (bulk RNA-seq)</div>
<div class="pub-badge"><a href="https://sung2021.github.io/samplePages/Daiichi_bulk_RNASeq/Daiichi_2024_bulk_30342_ElasticNet_featureSelection.v2.html" target="_blank">View Example →</a></div>
</div>
</div>
</div>

<div class="row">
<div class="col-md-4">
<div class="workflow-card">
<div class="pub-journal">Chromatin & Epigenomics</div>
<div class="pub-title">ChIP-seq, ATAC-seq, CUT&RUN pipelines — peak calling, motif analysis</div>
<div class="pub-badge"><a href="https://sung2021.github.io/samplePages/ChIP_seq.html" target="_blank">View Workflow →</a></div>
</div>
</div>

<div class="col-md-4">
<div class="workflow-card">
<div class="pub-journal">Omics Analytical Pipelines</div>
<div class="pub-title">Pipelines Collections; scRNAseq, RNA-seq, DNA-seq, ML</div>
<div class="pub-badge"><a href="https://sung2021.github.io/etc/General_workflow2.html" target="_blank">View Workflow →</a></div>
</div>
</div>

</div>

---

# Experience {#experience}

- Battelle (2025-Present) : Data Scientist II, Bioinformatics
- Dana Farber Cancer Institute (2022-2024) : Senior Bioinformatics Scientist 
- Hackensack Meridian Health (2021-2022) : Postdoctral fellow
- University of Michigan (2017-2020) : Postdoctral fellow

---

# GitHub Repositories {#repos}

```{r github-repos, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)

repos <- data.frame(
  Repository = c(
    '<a href="https://github.com/Sung2021/tle3-multiomics-NI2024" target="_blank">tle3-multiomics-NI2024</a>',
    '<a href="https://github.com/Sung2021/pubmed-local-rag" target="_blank">pubmed-local-rag</a>',
    '<a href="https://github.com/Sung2021/lung-tme-deconv-profiler" target="_blank">lung-tme-deconv-profiler</a>',
    '<a href="https://github.com/Sung2021/rnaseq_local_chat" target="_blank">rnaseq_local_chat</a>'
  ),
  Description = c(
    "Multi-omics analysis for the TLE3 study (Nature Immunology, 2024), integrating RNA-seq, ATAC-seq, and ChIP-seq.",
    "Local RAG pipeline for querying PubMed literature using LLMs — offline, context-aware search without API calls.",
    "Bulk RNA-seq deconvolution pipeline for tumor microenvironment (TME) profiling in lung cancer.",
    "Local LLM chat interface for interactive RNA-seq exploration — query DEG and pathway results in natural language."
  ),
  stringsAsFactors = FALSE
)

library(kableExtra)

kable(repos, format = "html", escape = FALSE, row.names = FALSE,
      col.names = c("Repository", "Description")) %>%
  kable_styling(bootstrap_options = c("bordered", "hover", "striped"),
                full_width = TRUE,
                font_size = 12)
```

---

# Contact {#contact}

- Email: sungrye@utexas.edu
- [GitHub](https://github.com/sung2021)
- [LinkedIn](https://www.linkedin.com/in/sung-rye-park-8a130098/)
