---
title: "Basic Github Usage"
subtitle: ""
date: "`r format(Sys.Date())`"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    toc_depth: 4
    code_folding: show
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=T, eval = F,
                      fig.align = "center", 
                      message=F, warning=F,
                      results = "markup",
                      error = TRUE,
                      highlight = TRUE,
                      prompt = FALSE,
                      tidy = FALSE)
```
<link rel="stylesheet" href="https://cdnjs.cloudflare.com/ajax/libs/font-awesome/6.5.0/css/all.min.css">

<style>
h1 { font-size: 1.75em; border-bottom: 2px solid #18bc9c; padding-bottom: 6px; margin-top: 40px; }
h2 { font-size: 1.4em; margin-top: 28px; }
h3 { font-size: 1.1em; }

body { font-size: 14.5px; color: #333; line-height: 1.7; }

.back-btn {
  display: inline-block;
  margin-bottom: 24px;
  padding: 6px 14px;
  background: #18bc9c;
  color: #fff !important;
  border-radius: 5px;
  font-size: 0.85em;
  text-decoration: none;
}
.back-btn:hover { background: #15a589; }

.info-box {
  background: #f4fbf9;
  border-left: 4px solid #18bc9c;
  border-radius: 0 6px 6px 0;
  padding: 12px 18px;
  margin: 16px 0 20px;
  font-size: 0.88em;
  color: #444;
}

pre {
  background: #f8f9fa;
  border: 1px solid #e3e6ea;
  border-radius: 5px;
  font-size: 0.87em;
}
</style>



```{r, echo=FALSE, eval=T}
library(dplyr)
library(ggplot2)
library(reshape)
```


<a href="https://sung2021.github.io/" class="back-btn"><i class="fa fa-arrow-left"></i> Back to Main Page</a>


### Step 1: Create a Local Repository
- **Create a local directory** for your repository:
  ```bash
  mkdir my_repository.io
  ```
- **Navigate into your directory**:
  ```bash
  cd path/my_repository.io
  ```

### Step 2: Initialize the Local Repository
- **Initialize Git** in your local directory:
  ```bash
  git init
  ```

### Step 3: Connect Your Local Repository to GitHub
- **Add a remote repository** URL to your local repository:
  ```bash
  git remote add origin https://github.com/Sung2021/my_repository.io
  ```

### Step 4: Add and Commit Files
- **Add files** to the staging area:
  ```bash
  git add . # Add all files
  git add filename # Add a specific file
  ```
- **Check the status** of your changes:
  ```bash
  git status
  ```
- **Commit your changes** with a message:
  ```bash
  git commit -m "commit message"
  ```

### Step 5: Push Changes to GitHub
- **Check which branch you are on**:
  ```bash
  git branch
  ```
- **Push your changes** to GitHub:
  ```bash
  git push origin main # Push to the main branch
  git push -f origin main # Force push to overwrite changes in the remote
  ```

### Step 6: Pull Latest Changes from GitHub
- **Pull changes** from the remote repository to your local repository:
  ```bash
  git pull origin main
  ```

---
### Password authentication using Token {.tabset}

#### Option 1: Using a Personal Access Token (PAT)

1. **Generate a Personal Access Token:**
   - Go to your GitHub account.
   - Navigate to Settings > Developer settings > Personal access tokens.
   - Click on "Generate new token".
   - Select the scopes or permissions you'd like to grant this token. For pushing to a repository, select `repo`.
   - Generate the token and make sure to copy it; you won’t be able to see it again once you navigate away from the page.
<br>

2. **Use the PAT instead of your password:**
   - When prompted for a password in the terminal, enter your personal access token instead of your GitHub password.

#### Option 2: Using SSH Keys

1. **Generate SSH Key Pair:**
   - Open your terminal and run `ssh-keygen -t ed25519 -C "your_email@example.com"`.
   - Follow the prompts to create your keys, ideally without a passphrase for automation.
   - This will create a public and private key pair. The public key ends in `.pub` and is safe to share.
<br>
2. **Add your SSH public key to GitHub:**
   - Navigate to your GitHub account.
   - Go to Settings > SSH and GPG keys.
   - Click on "New SSH key", give it a title, and paste your public key.
<br>
3. **Switch your remote URL from HTTPS to SSH:**
   - Check your current remote URL: `git remote -v`.
   - Change it to SSH: `git remote set-url origin git@github.com:username/repository.git`.
<br>
4. **Push again using SSH:**
   - Simply run your push command again: `git push origin main`.