---
title: "Data Processing in Linux"
subtitle: ""
date: "`r format(Sys.Date())`"
output:  
  rmdformats::robobook: 
    code_folding: show 
    number_sections: FALSE
    toc_depth: 6
    toc_float: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=T, eval = F,
                      fig.align = "center", 
                      message=F, warning=F,
                      results = "markup",
                      error = TRUE,
                      highlight = TRUE,
                      prompt = FALSE,
                      tidy = FALSE)
```


```{r, echo=FALSE, eval=T}
library(dplyr)
library(ggplot2)
library(reshape)
```


<a href="https://sung2021.github.io/">Back to Main Page </a>  

<br>

### Step 1: Create a Local Repository
- **Create a local directory** for your repository:
  ```bash
  mkdir my_repository.io
  ```
- **Navigate into your directory**:
  ```bash
  cd path/my_repository.io
  ```

### Step 2: Initialize the Local Repository
- **Initialize Git** in your local directory:
  ```bash
  git init
  ```

### Step 3: Connect Your Local Repository to GitHub
- **Add a remote repository** URL to your local repository:
  ```bash
  git remote add origin https://github.com/Sung2021/my_repository.io
  ```

### Step 4: Add and Commit Files
- **Add files** to the staging area:
  ```bash
  git add . # Add all files
  git add filename # Add a specific file
  ```
- **Check the status** of your changes:
  ```bash
  git status
  ```
- **Commit your changes** with a message:
  ```bash
  git commit -m "commit message"
  ```

### Step 5: Push Changes to GitHub
- **Check which branch you are on**:
  ```bash
  git branch
  ```
- **Push your changes** to GitHub:
  ```bash
  git push origin main # Push to the main branch
  git push -f origin main # Force push to overwrite changes in the remote
  ```

### Step 6: Pull Latest Changes from GitHub
- **Pull changes** from the remote repository to your local repository:
  ```bash
  git pull origin main
  ```

