---
title: "Combinational test of DGKi with PD-1 blocker"
subtitle: "scTCR-seq + scRNA-seq"
author: "Sung Rye Park"
date: "`r format(Sys.Date())`"
output:  
  rmdformats::robobook: 
  code_folding: hide 
  number_sections: FALSE
  toc_depth: 6
  toc_float: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=F, fig.align = "center", 
                      message=F, warning=F,
                      results = "markup",
                      error = TRUE,
                      highlight = TRUE,
                      prompt = FALSE,
                      tidy = FALSE)
```

# DGKi 

- DGKs are physiological regulators of T-cell development,differentiation, and function.   
- Inhibition of DGK may restore the T-cell activation program   
- Enhanced effects may be seen with combination therapy (Nivolumab + DGKi)   
 
<br>


```{r, echo=F}
library(dplyr)
library(ggplot2)
library(reshape)
library(Seurat)
library(ggthemes)
```

# Data  

Sample name: P*****  
Condition (1 control, 3 test conditions): CONTROL, CND1, CND2, COMBO  
Number of cells : 8096 cells     

```{r}
dir="~/Desktop/DF/DFCI_Paweletz/2023_Lauren_BMS_DGKi_23.03.07/"
obj.srt = readRDS(paste0(dir,'rds/P30276.8096cells.23.07.15.rds'))
```

## Initial Processing  

```{r, results='hide', eval=TRUE, echo=TRUE}
## perform default analysis
perform_default_analysis <- function(obj.srt, n_features = 2000, n_pcs = 20, 
                                     dims_for_neighbors = 1:10, 
                                     resolutions = c(0.2, 0.5), 
                                     umap_dims = 1:10) {
  # Step 1: Find variable features
  obj.srt <- FindVariableFeatures(obj.srt, 
                                  selection.method = 'vst', 
                                  nfeatures = n_features)
  
  # Step 2: Scale and normalize data
  all_genes <- rownames(obj.srt)
  obj.srt <- NormalizeData(obj.srt)
  obj.srt <- ScaleData(obj.srt, features = all_genes)
  
  # Step 3: Run PCA
  obj.srt <- RunPCA(obj.srt, 
                    features = VariableFeatures(object = obj.srt), npcs = n_pcs)
  
  # Step 4: Find neighbors
  obj.srt <- FindNeighbors(obj.srt, dims = dims_for_neighbors)
  
  # Step 5: Find clusters
  obj.srt <- FindClusters(obj.srt, resolution = resolutions)
  
  # Step 6: Run UMAP
  obj.srt <- RunUMAP(obj.srt, dims = umap_dims)
  
  # Return the Seurat object with analysis results
  return(obj.srt)
}

# apply
obj.srt <- perform_default_analysis(obj.srt)
```


## Numer of cells  
```{r}
obj.srt@meta.data  = obj.srt@meta.data %>% 
  mutate(condition = ifelse(sample == "C276", "CTRL", 
                            ifelse(sample == "D276","CND1",
                                   ifelse(sample =="N276", "CND2",
                                          ifelse(sample == "DN276", "COMB","NO")))))

obj.srt@meta.data$condition = factor(obj.srt@meta.data$condition, 
                                     levels = c("CTRL","CND1","CND2","COMB"))
obj.srt@meta.data$condition %>% table() %>% data.frame() %>% 
  DT::datatable(rownames = F, colnames = c("Condition","Number of Cells"))
```


## UMAP (sample) {.tabset}

### Colored by condition  

```{r}
wes = c("#f5c600","#d8460b","#9b4923","#007291")
```

```{r, fig.width=5, fig.height=4}
DimPlot(obj.srt, group.by = "condition", pt.size = 0.1, 
        cols = wes)
```


### Colored by each condition   
```{r}
umap_one_population = function(cond, color2="orangered4"){
  obj.srt@meta.data = obj.srt@meta.data %>% mutate(color=ifelse(condition == cond, cond,"other"))
  obj.srt$color = factor(obj.srt$color, levels = c(cond,"other"))
  
  df = obj.srt@reductions$umap@cell.embeddings %>% 
    data.frame(check.names = F) 
  df$color= obj.srt$color
  df$color = factor(df$color, levels = c(cond,"other"))
  
  # Plot with ggplot
  ggplot(df, aes(umap_1 ,umap_2 , color=color)) +  # Assume the columns are V1 and V2, replace with actual column names
    geom_point(data = df %>% filter(color == "other"), aes(color = color), size = 0.2, alpha = 0.6) +
    geom_point(data = df %>% filter(color == cond), aes(color = color), size = 0.2, alpha = 0.6) +
    scale_color_manual(values = c( "grey88",  color2)) +  # Replace 'cond' with the actual condition value
    theme_bw() +
    xlab("UMAP1") + ylab("UMAP2") +
    guides(colour = guide_legend(override.aes = list(size = 5))) +
    theme(legend.title = element_blank())
  
}

p1 =umap_one_population(cond = "CTRL", color2 = wes[1])
p2 =umap_one_population(cond = "CND1", color2 = wes[2])
p3 =umap_one_population(cond = "CND2", color2 = wes[3])
p4 =umap_one_population(cond = "COMB", color2 = wes[4])

cowplot::plot_grid(p1,p2,p3,p4, ncol = 2)
```

# Clustering {.tabset}   

## UMAP colored by cluster (two resolutions)   
```{r, fig.width=12, fig.height=5}
DimPlot(obj.srt, group.by = c("RNA_snn_res.0.2", "RNA_snn_res.0.5"), size=0.1)
```


## Distribution of cells by cluster and condition (res 0.1)   
```{r, fig.width=5, fig.height=4}
obj.srt@meta.data %>% ggplot(aes(RNA_snn_res.0.2, fill=condition)) + geom_bar(position = "fill") + scale_fill_manual(values = wes) + theme_bw()
```


## TABLE : Distribution of cells by cluster and condition (res 0.1)    
```{r}
obj.srt@meta.data %>% select(RNA_snn_res.0.2, condition) %>% table() %>% data.frame() %>% tidyr::spread(condition, Freq) %>% DT::datatable(rownames = F)
```
## HEATMAP : Distribution of cells by cluster and condition (res 0.1)    
```{r, fig.width=5, fig.height=4}
df = obj.srt@meta.data %>% select(RNA_snn_res.0.2, condition) %>% table() %>% data.frame() %>% tidyr::spread(condition, Freq) %>% select(2:5) 
for(i in 1:length(colnames(df))){
  df[,i]= df[,i]/sum(df[,i])*100
}

df %>% pheatmap::pheatmap(display_numbers = T, number_format = "%.0f", cluster_rows = F, cluster_cols = F, color= colorRampPalette(colors = c("white","brown"))(1000), main = "Number of cells (%)", border_color = "grey99")
```



## Distribution of cells by cluster and condition (res 0.5)   
```{r, fig.width=7, fig.height=4}
obj.srt@meta.data %>% ggplot(aes(RNA_snn_res.0.5, fill=condition)) + geom_bar(position = "fill") + scale_fill_manual(values = wes) + theme_bw()
```

## TABLE : Distribution of cells by cluster and condition (res 0.5)    
```{r}
obj.srt@meta.data %>% select(RNA_snn_res.0.5, condition) %>% table() %>% data.frame() %>% tidyr::spread(condition, Freq) %>% DT::datatable(rownames = F)
```


## HEATMAP : Distribution of cells by cluster and condition (res 0.5)    
```{r, fig.width=5, fig.height=4}
df = obj.srt@meta.data %>% select(RNA_snn_res.0.5, condition) %>% table() %>% data.frame() %>% tidyr::spread(condition, Freq) %>% select(2:5) 
for(i in 1:length(colnames(df))){
  df[,i]= df[,i]/sum(df[,i])*100
}

df %>% pheatmap::pheatmap(display_numbers = T, number_format = "%.0f", cluster_rows = F, cluster_cols = F, color= colorRampPalette(colors = c("white","brown"))(1000), main = "Number of cells (%)", border_color = "grey99")
```

<br><br>
# Clonotype Analysis     

# Clonotype Diversity {.tabset}  

```{r}
# Define function 
library(packcircles)
bubbleplot= function(data){
  data[,1] = stringr::str_replace(data[,1],"clonotype", "clon")
  
  # Generate the layout. This function return a dataframe with one line per bubble. 
  # It gives its center (x and y) and its radius, proportional of the value
  packing <- circleProgressiveLayout(data$Freq, sizetype='area')
  
  # We can add these packing information to the initial data frame
  data <- cbind(data, packing)
  colnames(data)[1] ='clonotype'
  dat.gg <- circleLayoutVertices(packing, npoints=50)
  
  p=ggplot() + 
    # Make the bubbles
    geom_polygon(data = dat.gg, aes(x, y, group = id, 
                                    fill=as.factor(id)), colour = "#424949", alpha = 0.6) +
    # General theme:
    theme_void() + 
    theme(legend.position="none") +
    coord_equal() +
    
    # Add text in the center of each bubble + control its size
    geom_text(data = data, aes(x, y, size=Freq, label = clonotype)) +
    scale_size_continuous(range = c(1,6)) +
    labs(title="Clonotype Diversity")
  return(p)
}

```

## CTRL clonotype diversity  
```{r}
C276= obj.srt@meta.data %>% filter(condition == "CTRL") %>% select(raw_clonotype_id) %>% table() %>% data.frame()
bubbleplot(data = C276)
```

## CND1 clonotype diversity  
```{r}
D276= obj.srt@meta.data %>% filter(condition == "CND1") %>% select(raw_clonotype_id) %>% table() %>% data.frame()
bubbleplot(data = D276)
```


## CND2 clonotype diversity  
```{r}
N276= obj.srt@meta.data %>% filter(condition == "CND2") %>% select(raw_clonotype_id) %>% table() %>% data.frame()
bubbleplot(data = N276)
```


## COMB clonotype diversity  
```{r}
DN276= obj.srt@meta.data %>% filter(condition == "COMB") %>% select(raw_clonotype_id) %>% table() %>% data.frame()
bubbleplot(data = DN276)
```



# scRepertoire  

Clonotype analysis using scRepertoire   
```{r}
library(scRepertoire)
# obj.srt@meta.data[1:3,]

l1 = obj.srt@meta.data %>% filter(condition=="CTRL") %>% select(14:43)
l2 = obj.srt@meta.data %>% filter(condition=="CND1") %>% select(14:43)
l3 = obj.srt@meta.data %>% filter(condition=="CND2") %>% select(14:43)
l4 = obj.srt@meta.data %>% filter(condition=="COMB") %>% select(14:43)


contig_list = list(l1=l1,l2=l2,l3=l3,l4=l4)

combined <- combineTCR(contig_list, samples =  unique(obj.srt@meta.data$condition), cells = "T-AB")
```


## Quantify Clonotypes
```{r}
quantContig(combined, cloneCall="gene+nt", scale = T)
```

## Unique clonotypes by TRA chain      
```{r}
quantContig(combined, cloneCall="gene+nt", chain = "TRA")
```

## Unique clonotypes by TRB chain      
```{r}
quantContig(combined, cloneCall="gene+nt", chain = "TRB")
```

# Clonotype Abundance   

```{r}
abundanceContig(combined, cloneCall = "gene", scale = F)
```


# Length of Clonotypes  

```{r}
lengthContig(combined, cloneCall="aa", chain = "both") 
```

```{r}
lengthContig(combined, cloneCall="nt", chain = "TRA") + ggtitle("TRA")
```


```{r}
lengthContig(combined, cloneCall="nt", chain = "TRB") + ggtitle("TRB")
```


# Compare Clonotypes  

```{r}
compareClonotypes(combined, 
                  numbers = 5, 
                  samples = c("CTRL","CND1","CND2","COMB"), 
                  cloneCall="aa", 
                  graph = "alluvial")
```


# Visualize Gene Usage   

```{r}
vizGenes(combined, gene = "V", 
         chain = "TRB", 
         plot = "bar", 
         order = "variance", 
         scale = TRUE)
```

```{r}
#Peripheral Blood
vizGenes(combined, 
         gene = "V", 
         chain = "TRB", 
         y.axis = "J", 
         plot = "heatmap", 
         scale = TRUE, 
         order = "gene")
```

# More Advanced Clonal Analysis   

```{r}
clonalHomeostasis(combined, cloneCall = "gene", 
                  cloneTypes = c(Rare = 1e-04, 
                                 Small = 0.001, 
                                 Medium = 0.01, 
                                 Large = 0.1, 
                                 Hyperexpanded = 1)) + ggtitle("cloneCall = gene")
```

```{r}
clonalHomeostasis(combined, cloneCall = "aa") + ggtitle("cloneCall = aa")
```

# Diversity Analysis 

```{r}
clonalDiversity(combined, 
                cloneCall = "gene", 
                group.by = "sample",
                n.boots = 100)
```


```{r}
# seurat <- combineExpression(combined, obj.srt, 
#                  cloneCall="gene",
#                  proportion = TRUE)
```


```{r, eval=FALSE}
# library(ggraph)
# 
# #No Identity filter
# clonalNetwork(seurat, 
#               reduction = "umap", 
#               identity = "sample",
#               filter.clones = NULL,
#               filter.identity = NULL,
#               cloneCall = "aa")
```





```{r}
# library(circlize)
# library(scales)
# 
# circles <- getCirclize(seurat, 
#                        group.by = "sample")
# 
# #Just assigning the normal colors to each cluster
# grid.cols <- scales::hue_pal()(length(unique(seurat@active.ident)))
# names(grid.cols) <- levels(seurat@active.ident)
# 
# #Graphing the chord diagram
# circlize::chordDiagram(circles,
#                        self.link = 1, 
#                       grid.col = grid.cols)
```

<br>

# Further Analysis (Specific cell type+clonotype)   

Cytotoxic Markers and associated clonotypes  

```{r, eval=TRUE, fig.width=6, fig.height=6}
# obj.srt@meta.data[1:3,]

df = obj.srt@meta.data %>% filter(cytotoxic == "cytotoxic") %>% 
  select(raw_clonotype_id) %>% table() %>% data.frame() %>% 
  arrange(desc(Freq))
df = df[1:10,]
# df[1:3,]

# cytotoxic T cell clonotypes 

obj.srt@meta.data %>% filter(cytotoxic == "cytotoxic") %>% 
  filter(raw_clonotype_id %in% df$raw_clonotype_id) %>% nrow()


df.anno = obj.srt@meta.data %>% filter(cytotoxic == "cytotoxic") %>% 
  filter(raw_clonotype_id %in% df$raw_clonotype_id)
df.anno= df.anno[,c("raw_clonotype_id","condition")] %>% arrange(desc(raw_clonotype_id))
obj.srt@meta.data %>% filter(cytotoxic == "cytotoxic") %>% 
  filter(raw_clonotype_id %in% df$raw_clonotype_id) %>% select(IFNG,GZMA,GZMB, NKG7) %>% 
  pheatmap::pheatmap(show_rownames = F, color = colorRampPalette(colors=c('ivory',"salmon"))(100),
                     annotation_row = df.anno, fontsize_col = 15)

```

```{r, fig.width=6, fig.height=6}

df.input = obj.srt@meta.data %>% filter(cytotoxic == "cytotoxic") %>% 
  filter(raw_clonotype_id %in% df$raw_clonotype_id) %>% select(IFNG,GZMA,GZMB, NKG7)
df.input[rownames(df.anno),] %>% 
  pheatmap::pheatmap(show_rownames = F, color = colorRampPalette(colors=c("white","salmon"))(100),
                     annotation_row = df.anno, cluster_rows = F)


```


