---
title: "RNA seq"
subtitle: "DEG to Kmeans clustering"
author: "Sung Rye Park"
date: "`r format(Sys.Date())`"
output:  
  rmdformats::robobook: 
    code_folding: hide 
    number_sections: FALSE
    toc_depth: 6
    toc_float: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=F, fig.align = "center", 
                      message=F, warning=F,
                      results = "markup",
                      error = TRUE,
                      highlight = TRUE,
                      prompt = FALSE,
                      tidy = FALSE)
```


# Load packages  
```{r, echo=TRUE}
library(dplyr)
library(ggplot2)
library(DESeq2)
library(edgeR)
library(reshape)
```


```{r, echo=FALSE, eval=TRUE}
# Read filtered count matrix 
count.raw = read.csv(paste0(dir,"rds/ITK_SYK_raw.count.filtered.24.04.12.csv"), row.names = 1)
count.raw = count.raw[,1:14]
count.mtx = count.raw[,c(1:4,7:9,10:13)]
colnames(count.mtx) = c(paste0("CND1_", 1:4), 
                        paste0("CND2_", 1:4),
                        paste0("CTRL_", 1:3))
tpm=read.csv(paste0(dir,"rds/tpm/RNA_seq.TPM.23.03.16.csv"), row.names = 1)
tpm = tpm[,1:14]
tpm = tpm[,c(1:4,7:9,10:13)]
colnames(tpm) = c(paste0("CND1_", 1:4), 
                        paste0("CND2_", 1:4),
                        paste0("CTRL_", 1:3))
```


# TPM table (downloadable)   
This is the example TPM value. (100 genes x 14 sample)

```{r}
# tpm %>% boxplot()

tpm[1:100,] %>% DT::datatable(extensions = "Buttons", 
                options = list(autoWidth = TRUE, 
                               fixedHeader = TRUE,
                               dom="Bfrtip", buttons=c("csv","excel")))
```



```{r, echo=FALSE}
zscore <- function(input.data = input.data){
  input.data.rowsums <- rowSums(input.data)
  input.data.mean <- rowMeans(input.data)
  input.data.sd <- matrixStats::rowSds(as.matrix(input.data))
  names(input.data.sd) <- rownames(input.data)
  zscore <- (input.data-input.data.mean)/input.data.sd
  return(zscore)
}

```


# PCA {.tabset} 
```{r}
se <- SummarizedExperiment(as.matrix(count.mtx), 
                           colData=DataFrame(sample=1:ncol(count.mtx)))
dds <- DESeqDataSet(se, ~ 1)
dds$sample = colnames(dds)
vsd <- vst(dds, blind=FALSE)
```

  
```{r}
pcaData <- DESeq2::plotPCA(vsd, intgroup = "sample", returnData = TRUE)
PCA_var=attr(pcaData, "percentVar")
pcaData$cond = substr(pcaData$sample,1,4)
```

## PCA plot   
```{r, fig.width=6.5, fig.height=5}
# Create a PCA plot
pcaData$cond = factor(pcaData$cond, levels = c("CTRL", "CND1", "CND2"))
ggplot(pcaData, aes(x = PC1, y = PC2, fill = cond)) +
  geom_point(size = 5, alpha = 0.6, shape = 21, color = "black", stroke = 0.5)  +
  ggrepel::geom_text_repel(aes(label=sample), color="grey6", size=3, hjust= -0.3, vjust=-0.3) +
  labs(x = paste("PC1: ", round(100 * PCA_var[1]), "% variance"),
       y = paste("PC2: ", round(100 * PCA_var[2]), "% variance")) +
  theme_bw() +
  theme(legend.title = element_blank())
```

## PCA plot by condition    
```{r, fig.width=6.5, fig.height=5}
# Create a PCA plot
ggplot(pcaData, aes(x = PC1, y = PC2, fill = cond)) +
  geom_point(size = 4, alpha = 0.6, shape = 21, color = "black", stroke = 0.5)  +
  ggrepel::geom_text_repel(aes(label=sample), color="grey6", size=3, hjust= -0.3, vjust=-0.3) +
  labs(x = paste("PC1: ", round(100 * PCA_var[1]), "% variance"),
       y = paste("PC2: ", round(100 * PCA_var[2]), "% variance")) +
  theme_bw() +
  theme(legend.title = element_blank()) +
  facet_wrap(.~cond, ncol = 2)
```

## PCA plot zoom in    
```{r, fig.width=6.5, fig.height=5}
# Create a PCA plot
ggplot(pcaData, aes(x = PC1, y = PC2, fill = cond)) +
  geom_point(size = 5, alpha = 0.6, shape = 21, color = "black", stroke = 0.5)  +
  ggrepel::geom_text_repel(aes(label=sample), color="grey6", size=3, hjust= -0.3, vjust=-0.3) +
  labs(x = paste("PC1: ", round(100 * PCA_var[1]), "% variance"),
       y = paste("PC2: ", round(100 * PCA_var[2]), "% variance")) +
  theme_bw() +
  theme(legend.title = element_blank()) +
  xlim(c(-15,-5)) + ylim(c(-2,2))
```



# DEG analysis {.tabset} 

```{r}
# Function 1 
# condition1 = CTRL 
DEG_out = function(count.raw= count.mtx, condition1, condition2){
  df1 = count.raw %>% select(contains(condition1)) 
  df2 = count.raw %>% select(contains(condition2)) 
  
  count.mtx = cbind(df1,df2)
  
  # Generate info table
  info <- data.frame(matrix(nrow = ncol(count.mtx), ncol = 2))
  colnames(info) <- c('sample', 'cond')
  info$sample <- colnames(count.mtx)
  info = info %>% mutate(cond = ifelse(sample %in% colnames(df1), condition1,condition2)) 
  info$cond = factor(info$cond, levels = c(condition1, condition2))
  # info
  
  # DESeq: DEG analysis 
  dds <- DESeqDataSetFromMatrix(count.mtx, info, ~ cond)
  dds <- DESeq(dds)
  res <- results(dds)
  
  # make DEG output a data frame for downstream analysis
  res <- data.frame(res)
  res <- res %>% filter(baseMean != 0)
  
  # p value == NA adjustment 
  res[is.na(res)] <- 1
  
  # res[is.na(res$pvalue),]
  return(res)
}


# Function 2 for volcanoplot  
DEG_volcanoplot = function(res= res ,pval=0.05, fc = 1.2, 
                           condition1=condition1, 
                           condition2=condition2){
  res = res %>% mutate(DE=ifelse(log2FoldChange >= log2(fc) & pvalue <= pval, 'UP',
                                 ifelse(log2FoldChange <= -log2(fc) & pvalue <= pval, 'DN','no_sig')))
  res$DE = factor(res$DE, levels = c('UP','DN','no_sig'))
  
  # Draw plot 
  res %>% 
    filter(pvalue !=1) %>% # remove the NA values 
    ggplot(aes(log2FoldChange, -log10(pvalue), color=DE)) + 
    geom_point(size=1, alpha=0.5) + 
    scale_color_manual(values = c('red','blue','grey')) +
    theme_classic() +
    geom_vline(xintercept = c(-log2(fc),log2(fc)), color='grey') +
    geom_hline(yintercept = -log10(0.05),color='grey') +
    guides(colour = guide_legend(override.aes = list(size=5))) +
    ggtitle(paste0(condition2, " / ", condition1)) +
    ggeasy::easy_center_title() ## to center title
  
}

# Function 3 
DEG_info = function(res, n=20, pval=0.05, fc = 1.2){
  res = res %>% mutate(DE=ifelse(log2FoldChange >= log2(fc) & pvalue <= pval, 'UP',
                                 ifelse(log2FoldChange <= -log2(fc) & pvalue <= pval, 'DN','no_sig')))
  res$DE = factor(res$DE, levels = c('UP','DN','no_sig'))
  
  df = res$DE %>% table() %>% data.frame()
  colnames(df) = c("DEG", "Number of genes")
  
  up = res %>% filter(DE == "UP") %>% arrange(desc(log2FoldChange)) %>% 
    rownames() %>% head(n) %>% paste0(collapse = ",")
  dn = res %>% filter(DE == "DN") %>% arrange(log2FoldChange) %>% 
    rownames() %>% head(n) %>% paste0(collapse = ",")
  
  df$top_genes_by_Log2FC =""
  df[1,3] = up
  df[2,3] = dn
  
  return(df)
}

DEG_table=function(res, pval=0.05, fc = 1.2){
  res = res %>% mutate(DE=ifelse(log2FoldChange >= log2(fc) & pvalue <= pval, 'UP',ifelse(log2FoldChange <= -log2(fc) & pvalue <= pval, 'DN','no_sig')))
  res$DE = factor(res$DE, levels = c('UP','DN','no_sig'))
  return(res)
}
```


## DEG CND1 vs CTRL  

1. DEG information table  
2. Volcanoplot  
3. DEG output. 

```{r, eval=TRUE}
DEG1 = DEG_out(count.raw = count.mtx, 
                   condition1 = "CTRL", condition2 = "CND1")

df = DEG_info(res = DEG1) 
df %>% DT::datatable(extensions = "Responsive")

DEG_volcanoplot(res = DEG1, condition1 = "CTRL", condition2= "CND1")

DEG1 %>% DEG_table() %>% DT::datatable(extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=10))
```



## DEG CND2 vs CTRL  

1. DEG information table  
2. Volcanoplot  
3. DEG output. 

```{r}

DEG2 = DEG_out(count.raw = count.mtx, 
                   condition1 = "CTRL", condition2 = "CND2")

df = DEG_info(res = DEG2) 
df %>% DT::datatable(extensions = "Responsive")

DEG_volcanoplot(res = DEG2, condition1 = "CTRL", condition2= "CND2")

DEG2 %>% DEG_table() %>% DT::datatable(extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=10))
```



## DEG CND2 vs CND1  

1. DEG information table  
2. Volcanoplot  
3. DEG output. 

```{r}

DEG3 = DEG_out(count.raw = count.mtx, 
                   condition1 = "CND1", condition2 = "CND2")

df = DEG_info(res = DEG3) 
df %>% DT::datatable(extensions = "Responsive")

DEG_volcanoplot(res = DEG3, condition1 = "CTRL", condition2= "CND2")

DEG3 %>% DEG_table() %>% DT::datatable(extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=10))
```



## Select the signicant DEGs  

From three comparisons   

* DEG1   
* DEG2  
* DEG3  

```{r}
significant_genes = function(res){
  pval=0.05
  fc = 1.2
  res = res %>% mutate(DE=ifelse(log2FoldChange >= log2(fc) & pvalue <= pval, 'UP',
                                 ifelse(log2FoldChange <= -log2(fc) & pvalue <= pval, 'DN','no_sig')))
  res$DE = factor(res$DE, levels = c('UP','DN','no_sig'))
  genes = res %>% filter(DE != 'no_sig') %>% rownames()
  return(genes)
}

```


```{r}
# Significant genes from each DEG outputs 
set1 = significant_genes(res = DEG1)
set2 = significant_genes(res = DEG2)
set3 = significant_genes(res = DEG3)

# Merge all outs 
significant_genes_all = union(set1,union(set2,set3))

# Check the number of genes 
#significant_genes_all %>% length()
```

```{r}
significant_genes_all[230:250]
```
<br><br>


# Kmeans clustering {.tabset}  

Significant genes X samples matrix was used for kmeans clustering to identify any patterns across data.  
Clustering method: Kmeans  
Adjustable parameter : k (number of clusters)   

```{r}
############## kmeans clustering #####################
## zscore function 
zscore <- function(input.data = input.data){
  input.data.rowsums <- rowSums(input.data)
  input.data.mean <- rowMeans(input.data)
  input.data.sd <- matrixStats::rowSds(as.matrix(input.data))
  names(input.data.sd) <- rownames(input.data)
  zscore <- (input.data-input.data.mean)/input.data.sd
  return(zscore)
}

# Create input data with zscaled 
input.data <- zscore(tpm[significant_genes_all,])
input.data <- na.omit(input.data)



# Perform kmeans clustering 
kmeans_clustering = function(k=3, input.data= input.data){
  set.seed(1234)
  fitout <- kmeans(input.data, centers = k, nstart = k)
  fitout.cluster <- fitout$cluster
  # table(fitout.cluster)
  df = fitout.cluster %>% as.data.frame()
  colnames(df) = "cluster"
  rownames(df) = names(fitout.cluster)
  df = df %>% arrange(cluster)
  df$cluster = paste0("Cluster", df$cluster)
  df$cluster = factor(df$cluster)
  return(df)
}

df3 = kmeans_clustering(k = 3, input.data = input.data)
df4 = kmeans_clustering(k = 4, input.data = input.data)
df5 = kmeans_clustering(k = 5, input.data = input.data)
df6 = kmeans_clustering(k = 6, input.data = input.data)

# Draw heatmap 
kmeans_heatmap = function(input.df, input.data){
  df.tmp = input.df$cluster %>% table() %>% data.frame()
  my.color=c(colorRampPalette(colors = c("royalblue","white"))(80),
             colorRampPalette(colors = c("white","indianred2"))(90))
  
  p =input.data[rownames(input.df),] %>% 
    pheatmap::pheatmap(cluster_cols = F, cluster_rows = F,
                       show_rownames = F,
                       annotation_row = input.df,
                       col = my.color,
                       fontsize_row = 6,fontsize_col = 10, 
                       border_color = 'NA', 
                       gaps_col = c(4,8), 
                       gaps_row = cumsum(df.tmp$Freq))
  print(p)
}
```


## K = 3 
```{r}
kmeans_heatmap(input.df = df3, input.data = input.data)

rows = intersect(rownames(df3), rownames(tpm))

cbind(df3[rows,],tpm[rows,]) %>% data.frame(check.names = F) %>% 
  DT::datatable(rownames = T, caption = "Clsuter information with TPM value", 
                colnames = c("cluster", colnames(tpm)))
```


## K = 4 
```{r}

kmeans_heatmap(input.df = df4, input.data = input.data)

rows = intersect(rownames(df4), rownames(tpm))

cbind(df4[rows,],tpm[rows,]) %>% data.frame(check.names = F) %>% 
  DT::datatable(rownames = T, caption = "Clsuter information with TPM value", 
                colnames = c("cluster", colnames(tpm)))
```

## K = 5 
```{r}

kmeans_heatmap(input.df = df5, input.data = input.data)

rows = intersect(rownames(df5), rownames(tpm))

cbind(df5[rows,],tpm[rows,]) %>% data.frame(check.names = F) %>% 
  DT::datatable(rownames = T, caption = "Clsuter information with TPM value", 
                colnames = c("cluster", colnames(tpm)))
```

## K = 6 
```{r}

kmeans_heatmap(input.df = df6, input.data = input.data)

rows = intersect(rownames(df6), rownames(tpm))

cbind(df6[rows,],tpm[rows,]) %>% data.frame(check.names = F) %>% 
  DT::datatable(rownames = T, caption = "Clsuter information with TPM value", 
                colnames = c("cluster", colnames(tpm)))
```


# Selected pathway analysis  {.tabset} 


```{r, fig.width=5, fig.height=5}
# enrichplot::gseaplot2 function 
perform_GSEA <- function(res, ref, pvalueCutoff = 1) {
  ranking <- function(res) {
    # Check the name of log2fc related 
    if ("avg_log2FC" %in% names(res)) {
      df <- res$avg_log2FC
    } else if ("log2FoldChange" %in% names(res)) {
      df <- res$log2FoldChange
    } else {
      stop("Neither avg_log2FC nor log2FoldChange columns found in the data frame.")
    }
    names(df) <- rownames(res)
    df <- sort(df, decreasing = TRUE)
    return(df)
  }
  
  ranked.res <- ranking(res)
  
  x <- clusterProfiler::GSEA(geneList = ranked.res,
                             TERM2GENE = ref,
                             pvalueCutoff = pvalueCutoff,
                             pAdjustMethod = "BH",
                             verbose = TRUE,
                             seed = TRUE)
  return(x)
}

enrichplot_leadingGenes_list = function(df, tpm){
  # Filter the dataframe such that only rows where the gene column values are in the row names of tpm are retained
  df = df[df$gene %in% rownames(tpm), ]
  
  genesUp = df %>% filter(position==1) %>% select(gene) %>% pull() %>% head(10)
  genesDN = df %>% filter(position==1) %>% arrange(desc(x)) %>% select(gene) %>% pull() %>% head(10)
  # Check the number of rows in the filtered data
  n = nrow(df)
  # Select the top 10 genes (use all data if there are less than 10 rows)
  # Create a dataframe and set up the structure
  df1 = data.frame(gene = c(genesUp, genesDN))
  # Assign "UP" and "DOWN" labels to the leading genes
  leading_at = c(rep("UP", length(genesUp)), rep("DOWN", length(genesDN)))
  df1$leading_at = leading_at
  # Return the dataframe
  return(df1)
}
enrichplot_leadingGenes_heatmap = function(df,red=50, blue=20, id=id){
  df = df[rownames(df)[df$gene %in% rownames(tpm)],]
  genesUp = df %>% filter(position==1) %>% select(gene) %>% pull() %>% head(10)
  genesDN = df %>% filter(position==1) %>% arrange(desc(x)) %>% select(gene) %>% pull() %>% head(10)
  
  zscore <- function(input.data = input.data){
    input.data.rowsums <- rowSums(input.data)
    input.data.mean <- rowMeans(input.data)
    input.data.sd <- matrixStats::rowSds(as.matrix(input.data))
    names(input.data.sd) <- rownames(input.data)
    zscore <- (input.data-input.data.mean)/input.data.sd
    return(zscore)
  }
  
  my.colors <- c(colorRampPalette(colors = c("navy","white"))(blue), 
                 colorRampPalette(colors = c("white","red"))(red))
  p = zscore(tpm[c(genesUp,genesDN),]) %>% 
    pheatmap::pheatmap(color = my.colors, cluster_cols = F, cluster_rows = F,
                       border_color = "grey3", 
                       main = id, fontsize_row = 8)
  
  print(p)
}


```


## IL6_JAK_STAT3_SIGNALING    


```{r}
# Input data part
id = "HALLMARK_IL6_JAK_STAT3_SIGNALING"
res = DEG1
```


```{r, fig.width=5, fig.height=4}

gsea.res = perform_GSEA(res = res, ref = hallmark, pvalueCutoff = 1)
enrichplot::gseaplot2(gsea.res, geneSetID = id, title = id)

# Getting leading gene information
gsea.res2 = enrichplot::gseaplot2(gsea.res, geneSetID = id, title = id)

# gsea.res2[[1]] # figure top
# gsea.res2[[2]] # figure middle
# gsea.res2[[2]] # figure bottom

df = gsea.res2[[1]]$data

enrichplot_leadingGenes_list(df=df, tpm=tpm) %>% DT::datatable()
enrichplot_leadingGenes_heatmap(df=df, red=40, blue=30, id=id)

```


## INFLAMMATORY RESPONSE   

```{r}
# Input data part
id = "HALLMARK_INFLAMMATORY_RESPONSE"
res = DEG2
```


```{r, fig.width=5, fig.height=4}

gsea.res = perform_GSEA(res = res, ref = hallmark, pvalueCutoff = 1)
enrichplot::gseaplot2(gsea.res, geneSetID = id, title = id)

# Getting leading gene information
gsea.res2 = enrichplot::gseaplot2(gsea.res, geneSetID = id, title = id)

# gsea.res2[[1]] # figure top
# gsea.res2[[2]] # figure middle
# gsea.res2[[2]] # figure bottom

df = gsea.res2[[1]]$data

enrichplot_leadingGenes_list(df=df, tpm=tpm) %>% DT::datatable()
enrichplot_leadingGenes_heatmap(df=df, red=50, blue=35, id=id)

```



# Next work  


Any application to reduce high variance across samples.  





<br><br>
