---
title: "Sample plots"
subtitle: ""
date: "`r format(Sys.Date())`"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    toc_depth: 4
    code_folding: show
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=T, fig.align = "center", 
                      message=F, warning=F,
                      results = "markup",
                      error = TRUE,
                      highlight = TRUE,
                      prompt = FALSE,
                      tidy = FALSE)
```
<link rel="stylesheet" href="https://cdnjs.cloudflare.com/ajax/libs/font-awesome/6.5.0/css/all.min.css">

<style>
h1 { font-size: 1.75em; border-bottom: 2px solid #18bc9c; padding-bottom: 6px; margin-top: 40px; }
h2 { font-size: 1.4em; margin-top: 28px; }
h3 { font-size: 1.1em; }

body { font-size: 14.5px; color: #333; line-height: 1.7; }

.back-btn {
  display: inline-block;
  margin-bottom: 24px;
  padding: 6px 14px;
  background: #18bc9c;
  color: #fff !important;
  border-radius: 5px;
  font-size: 0.85em;
  text-decoration: none;
}
.back-btn:hover { background: #15a589; }

.info-box {
  background: #f4fbf9;
  border-left: 4px solid #18bc9c;
  border-radius: 0 6px 6px 0;
  padding: 12px 18px;
  margin: 16px 0 20px;
  font-size: 0.88em;
  color: #444;
}

pre {
  background: #f8f9fa;
  border: 1px solid #e3e6ea;
  border-radius: 5px;
  font-size: 0.87em;
}
</style>



```{r, echo=FALSE}
library(dplyr)
library(ggplot2)
library(reshape)
library(dplyr)
```


<a href="https://sung2021.github.io/" class="back-btn"><i class="fa fa-arrow-left"></i> Back to Main Page</a>
  




# Sunburst graph plot  

Let's just learn the concept of sunburst graph
```{r}
library(plotly)
plot_ly(labels = c("Epcam+", 
                   "Trop2+", "Trop2-", 
                   "Trop2+EGFR+", "Trop2+EGFR-", "Trop2-EGFR+", "Trop2-EGFR-", 
                   "Trop2+EGFR+HER2+","Trop2+EGFR+HER2-","Trop2-EGFR+HER2+","Trop2-EGFR+HER2-"),
        parents = c("", 
                    "Epcam+", "Epcam+", 
                    "Trop2+", "Trop2+", "Trop2-", "Trop2-", 
                    "Trop2+EGFR+","Trop2+EGFR+","Trop2-EGFR-","Trop2-EGFR-"),
        values = c(46, 37, 9, 14, 23, 5, 4, 0, 14, 4,1), 
        type = 'sunburst')

```



<br><br>
More example plots will be updated.  

<br><br>

