---
title: "Genomic Data Processing"
subtitle: "ChIP-seq · ATAC-seq · CUT&RUN — Peak Calling, Normalization, R Analysis"
date: "`r format(Sys.Date())`"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    toc_depth: 4
    code_folding: none
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=T, eval = F,
                      fig.align = "center", 
                      message=F, warning=F,
                      results = "markup",
                      error = TRUE,
                      highlight = TRUE,
                      prompt = FALSE,
                      tidy = FALSE)
```

<link rel="stylesheet" href="https://cdnjs.cloudflare.com/ajax/libs/font-awesome/6.5.0/css/all.min.css">

<style>
h1 { font-size: 1.75em; border-bottom: 2px solid #18bc9c; padding-bottom: 6px; margin-top: 40px; }
h2 { font-size: 1.4em; margin-top: 28px; }
h3 { font-size: 1.1em; }

body { font-size: 14.5px; color: #333; line-height: 1.7; }

.back-btn {
  display: inline-block;
  margin-bottom: 24px;
  padding: 6px 14px;
  background: #18bc9c;
  color: #fff !important;
  border-radius: 5px;
  font-size: 0.85em;
  text-decoration: none;
}
.back-btn:hover { background: #15a589; }

.info-box {
  background: #f4fbf9;
  border-left: 4px solid #18bc9c;
  border-radius: 0 6px 6px 0;
  padding: 12px 18px;
  margin: 16px 0 20px;
  font-size: 0.88em;
  color: #444;
}

pre {
  background: #f8f9fa;
  border: 1px solid #e3e6ea;
  border-radius: 5px;
  font-size: 0.87em;
}
</style>

```{r, echo=FALSE, eval=T}
library(dplyr)
library(ggplot2)
library(reshape)
```

<a href="https://sung2021.github.io/" class="back-btn"><i class="fa fa-arrow-left"></i> Back to Main Page</a>

* Peak Calling 
* Normalization
* Peak analysis in R

# Peak Calling by MACS2 {.tabset} 

## Peak Calling of individual bam file {.tabset}

Input files:

* sample.bam 
* control.bam

### Script

```{bash}
nano run_peak_calling.sh

#!/bin/bash

# Check if required input BAM files are provided
if [ -z "$1" ] || [ -z "$2" ]; then
  echo "Usage: $0 <input_bam> <control_bam>"
  exit 1
fi

# BAM file (sample)
input_bam=$1

# Control input BAM file (background)
common_input=$2

# Output directory
output_dir="macs2_out"

# Ensure the output directory exists
mkdir -p "$output_dir"

# Extract sample name from the input BAM file
sample_name=$(basename "$input_bam")

# Run MACS2 peak calling
macs2 callpeak \
    -t "$input_bam" \
    -c "$common_input" \
    -f BAMPE \
    -g mm \
    -n "$sample_name" \
    -B \
    -q 0.05 \
    --outdir "$output_dir"

# Check exit status
if [ $? -ne 0 ]; then
  echo "Error: MACS2 peak calling failed."
  exit 1
fi

echo "MACS2 peak calling completed successfully. Results are in

```

### Run  (Example)
```{bash}
chmod +x run_peak_calling.sh

bash run_peak_calling.sh /path/to/sample.bam /path/to/control.bam
./run_peak_calling.sh DMSO_H3K27me3.merged.bam /path/to/control.bam
./run_peak_calling.sh Treat1_H3K27me3.merged.bam /path/to/control.bam
./run_peak_calling.sh Treat2_H3K27me3.merged.bam /path/to/control.bam
```



## Peak Calling putting a folder as input {.tabset}

### Script
```{bash}
nano macs2_callpeaks.sh

#!/bin/bash

# Check if the input folder is provided
if [ -z "$1" ]; then
  echo "Usage: $0 <input_folder>"
  exit 1
fi

# Input folder containing BAM files
input_folder=$1

# Control input BAM files (glob pattern inside the folder)
common_input="${input_folder}/*_H3K27.bam"

# Output directory
output_dir="macs2_out"

# Ensure the output directory exists
mkdir -p "$output_dir"

# Sample peak calling
for sample_bam in "${input_folder}"/*.merged.bam; do
    sample=$(basename "$sample_bam" .merged.bam)
    macs2 callpeak \
        -t "$sample_bam" \
        -c "$common_input" \
        -f BAMPE \
        -g mm \
        -n "$sample" \
        -B \
        -q 0.05 \
        --outdir "$output_dir"
done

echo "MACS2 peak calling completed. Results are in '$output_dir'."
```

### Run
```{bash}
chmod +x macs2_callpeaks.sh
./macs2_callpeaks.sh /path/to/input_folder
```

<br>

## Broad Peak option {.tabset}  

Broad Peak for Extended Regions:  
Focuses on large, spread-out areas suited for markers such as H3K27me3 or H3K36me3. 

### Script 

--broad-cutoff 0.1  (FDR 10%)   

```{bash}
#!/bin/bash

# BAM file
input_bam=$1

# Control input  
common_input="../input.bam" # My input 

# Output directory 
output_dir="macs2_out"

# sample name 
sample_name=$(basename $input_bam .merged.bam)

# MACS2 run for broad peaks
macs2 callpeak \
    -t $input_bam \
    -c $common_input \
    -f BAMPE \
    -g mm \
    -n $sample_name \
    -B \
    --broad \
    -q 0.05 \
    --broad-cutoff 0.1 \
    --outdir $output_dir



```



### Run
```{bash}
./run_broadpeak_calling.sh input.bam
```


<br><hr><br>

# Peak filtering   

## Narrowpeak column information  

* Check signalValue column 

```{r,echo=F, eval=T}
library(data.table)
narrowPeak_columns <- data.table(
  Column = c("chrom", "chromStart", "chromEnd", "name", "score", "strand", 
             "signalValue", "pValue", "qValue", "peak"),
  Description = c("Name of the chromosome.",
                  "Start position of the peak (0-based).",
                  "End position of the peak (1-based).",
                  "Name of the peak.",
                  "Score of the peak, the significance of the peak.",
                  "The DNA strand.",
                  "Signal value in the peak region, the intensity or density of the peak.",
                  "P-value.",
                  "Adjusted p-value.",
                  "The peak summit relative to chromStart, the highest point within the peak.")
)

narrowPeak_columns %>% DT::datatable(rownames = T)
```

## Filtering and Sorting  

* Filtering by signalValue if necessary.   
* Sorting bed:  
  * First order by chromosome names using -k1,1 and then by start positions numerically with -k2,2n.  
  * This sorts entries by chromosome and then by position within each chromosome.  
<br><br>

<hr>

# Normalization of peaks

Bed Coverage calculation from reference peak (unionpeak) information   

The adv/disadv of processing bam files using a standardized peak set (union peaks) as a reference   

  - Standardizes peak across different samples
  - Reduced Bias
  - Efficiency in the downstream analysis
  - Risk of overgeneralization (e.g. overlooking sample/condition-specific peaks like peak length)


## bedtools coverage  

```{bash}
#!/bin/bash

# Assign BAM file
bamFile=$1

# Assign reference Peak File
unionPeakFile=$2

# Run bedtools coverage
bedtools coverage -a "$unionPeakFile" -b "$bamFile" > "${bamFile%.bam}.cov.bed"

echo "Coverage calculation for $bamFile completed."


./bamcoverage_single.sh path/to/your.bam path/to/unionpeak.bed

```


# Peak analysis in R

## Read files in R   
```{r}
library(data.table)

# Getting files with .cov.bed 
cov_files <- list.files(pattern = "\\.cov\\.bed$")

# Import and save 
cov_data_list <- lapply(cov_files, function(file) {
  fread(file)
})

# assign names 
names(cov_data_list) <- cov_files
```


## Coverage bed file example  

```{r, echo=F, eval=TRUE}
# Loading necessary library
library(data.table)

# Create a sample data frame that might resemble a coverage file output
coverage_data <- data.table(
  chrom = c("chr1", "chr1", "chr2", "chr2", "chr3"),
  start = c(100, 500, 200, 800, 150),
  end = c(200, 600, 300, 900, 250),
  num_reads = c(25, 30, 5, 20, 15),  # Example read counts
  coverageFraction = c(1, 0.95, 0.20, 1, 0.85)  # Example mean coverage depth
)

# Print the data frame to see its content
print(coverage_data)
```
<br><br>
Modify coverage bed files based on the purpose in the downstream analysis.  

<br><br>
