---
title: "scTCR-seq analysis by scRepertoire"
subtitle: ""
author: "Sung Rye Park"
date: "`r format(Sys.Date())`"
output:  
  rmdformats::robobook: 
  code_folding: hide 
  number_sections: FALSE
  toc_depth: 6
  toc_float: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=F, fig.align = "center", 
                      message=F, warning=F,
                      results = "markup",
                      error = TRUE,
                      highlight = TRUE,
                      prompt = FALSE,
                      tidy = FALSE)
```



```{r, echo=F}
library(dplyr)
library(ggplot2)
library(reshape)
library(Seurat)
library(ggthemes)
```

# Data  

Sample name: P*****  
Condition (1 control, 3 test conditions): CONTROL, CND1, CND2, COMBO  
Number of cells : ~9000 cells     

```{r}
dir="~/Desktop/DF/DFCI_Paweletz/2023_Lauren_BMS_DGKi_23.03.07/"
obj.srt = readRDS(paste0(dir,'rds/P30276.8096cells.23.07.15.rds'))
```

 

```{r, results='hide', eval=TRUE, echo=FALSE}
## perform default analysis
perform_default_analysis <- function(obj.srt, n_features = 2000, n_pcs = 20, 
                                     dims_for_neighbors = 1:10, 
                                     resolutions = c(0.2, 0.5), 
                                     umap_dims = 1:10) {
  # Step 1: Find variable features
  obj.srt <- FindVariableFeatures(obj.srt, 
                                  selection.method = 'vst', 
                                  nfeatures = n_features)
  
  # Step 2: Scale and normalize data
  all_genes <- rownames(obj.srt)
  obj.srt <- NormalizeData(obj.srt)
  obj.srt <- ScaleData(obj.srt, features = all_genes)
  
  # Step 3: Run PCA
  obj.srt <- RunPCA(obj.srt, 
                    features = VariableFeatures(object = obj.srt), npcs = n_pcs)
  
  # Step 4: Find neighbors
  obj.srt <- FindNeighbors(obj.srt, dims = dims_for_neighbors)
  
  # Step 5: Find clusters
  obj.srt <- FindClusters(obj.srt, resolution = resolutions)
  
  # Step 6: Run UMAP
  obj.srt <- RunUMAP(obj.srt, dims = umap_dims)
  
  # Return the Seurat object with analysis results
  return(obj.srt)
}

# apply
obj.srt <- perform_default_analysis(obj.srt)
```



```{r, echo=FALSE}
obj.srt@meta.data  = obj.srt@meta.data %>% 
  mutate(condition = ifelse(sample == "C276", "CTRL", 
                            ifelse(sample == "D276","CND1",
                                   ifelse(sample =="N276", "CND2",
                                          ifelse(sample == "DN276", "COMB","NO")))))
```

# scRepertoire  

Clonotype analysis using scRepertoire package   

Analysis below represents the basic analysis using the package.  
```{r}
library(scRepertoire)
# obj.srt@meta.data[1:3,]

l1 = obj.srt@meta.data %>% filter(condition=="CTRL") %>% select(14:43)
l2 = obj.srt@meta.data %>% filter(condition=="CND1") %>% select(14:43)
l3 = obj.srt@meta.data %>% filter(condition=="CND2") %>% select(14:43)
l4 = obj.srt@meta.data %>% filter(condition=="COMB") %>% select(14:43)


contig_list = list(l1=l1,l2=l2,l3=l3,l4=l4)

combined <- combineTCR(contig_list, samples =  unique(obj.srt@meta.data$condition))
# combined <- combineTCR(contig_list, samples =  unique(obj.srt@meta.data$condition), cells = "T-AB")
```


<!-- ## Quantify Clonotypes -->
<!-- ```{r, fig.width=5, fig.height=4} -->
<!-- quantContig(combined, cloneCall="gene+nt", scale = T) -->
<!-- ``` -->

<!-- ## Unique clonotypes by TRA chain       -->
<!-- ```{r, fig.width=5, fig.height=4} -->
<!-- quantContig(combined, cloneCall="gene+nt", chain = "TRA") -->
<!-- ``` -->

<!-- ## Unique clonotypes by TRB chain       -->
<!-- ```{r, fig.width=5, fig.height=4} -->
<!-- quantContig(combined, cloneCall="gene+nt", chain = "TRB") -->
<!-- ``` -->

## Numbers of unique clones   

```{r, fig.width=6, fig.height=4}
clonalQuant(combined, 
            cloneCall="strict", 
            chain = "both", 
            scale = TRUE)
```


## clonalAbundance {.tabset}   

### No scale
```{r, fig.width=6, fig.height=4}
clonalAbundance(combined, 
                cloneCall = "gene", 
                scale = FALSE)
```
### Scaled
```{r, fig.width=6, fig.height=4}
clonalAbundance(combined, 
                cloneCall = "gene", 
                scale = TRUE)
```


```{r, fig.width=5, fig.height=4}
abundanceContig(combined, cloneCall = "gene", scale = F)
```


## Length of Clonotypes  

```{r, fig.width=6, fig.height=4}
lengthContig(combined, cloneCall="aa", chain = "both") 
```

```{r, fig.width=6, fig.height=4}
lengthContig(combined, cloneCall="nt", chain = "TRA") + ggtitle("TRA")
```


```{r, fig.width=6, fig.height=4}
lengthContig(combined, cloneCall="nt", chain = "TRB") + ggtitle("TRB")
```


## Compare Clonotypes  

```{r, fig.width=7, fig.height=4}
compareClonotypes(combined, 
                  numbers = 5, 
                  samples = c("CTRL","CND1","CND2","COMB"), 
                  cloneCall="aa", 
                  graph = "alluvial")
```


## Visualize Gene Usage   

```{r, fig.width=8, fig.height=6}
vizGenes(combined, gene = "V", 
         chain = "TRB", 
         plot = "bar", 
         order = "variance", 
         scale = TRUE)
```

```{r}
#Peripheral Blood
vizGenes(combined, 
         gene = "V", 
         chain = "TRB", 
         y.axis = "J", 
         plot = "heatmap", 
         scale = TRUE, 
         order = "gene")
```

## More Advanced Clonal Analysis   

```{r, fig.width=5, fig.height=4}
clonalHomeostasis(combined, cloneCall = "gene", 
                  cloneTypes = c(Rare = 1e-04, 
                                 Small = 0.001, 
                                 Medium = 0.01, 
                                 Large = 0.1, 
                                 Hyperexpanded = 1)) + ggtitle("cloneCall = gene")
```

```{r, fig.width=5, fig.height=4}
clonalHomeostasis(combined, cloneCall = "aa") + ggtitle("cloneCall = aa")
```

## Diversity Analysis 

```{r, fig.width=5, fig.height=4}
clonalDiversity(combined, 
                cloneCall = "gene", 
                group.by = "sample",
                n.boots = 100)
```


```{r}
# seurat <- combineExpression(combined, obj.srt, 
#                  cloneCall="gene",
#                  proportion = TRUE)
```


```{r, eval=FALSE}
# library(ggraph)
# 
# #No Identity filter
# clonalNetwork(seurat, 
#               reduction = "umap", 
#               identity = "sample",
#               filter.clones = NULL,
#               filter.identity = NULL,
#               cloneCall = "aa")
```





```{r}
# library(circlize)
# library(scales)
# 
# circles <- getCirclize(seurat, 
#                        group.by = "sample")
# 
# #Just assigning the normal colors to each cluster
# grid.cols <- scales::hue_pal()(length(unique(seurat@active.ident)))
# names(grid.cols) <- levels(seurat@active.ident)
# 
# #Graphing the chord diagram
# circlize::chordDiagram(circles,
#                        self.link = 1, 
#                       grid.col = grid.cols)
```


<hr>

<br><br>
