---
title: "scRNA-seq Analysis Workflow"
subtitle: "Seurat v5 · Harmony · scDblFinder · Monocle3"
date: "`r format(Sys.Date())`"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    toc_depth: 4
    code_folding: none
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=T, fig.align = "center", eval = F,
                      message=F, warning=F,
                      results = "markup",
                      error = TRUE,
                      highlight = TRUE,
                      prompt = FALSE,
                      tidy = FALSE)
```

<link rel="stylesheet" href="https://cdnjs.cloudflare.com/ajax/libs/font-awesome/6.5.0/css/all.min.css">

<style>
h1 { font-size: 1.75em; border-bottom: 2px solid #18bc9c; padding-bottom: 6px; margin-top: 40px; }
h2 { font-size: 1.4em; margin-top: 28px; }
h3 { font-size: 1.1em; }

body { font-size: 14.5px; color: #333; line-height: 1.7; }

.back-btn {
  display: inline-block;
  margin-bottom: 24px;
  padding: 6px 14px;
  background: #18bc9c;
  color: #fff !important;
  border-radius: 5px;
  font-size: 0.85em;
  text-decoration: none;
}
.back-btn:hover { background: #15a589; }

.info-box {
  background: #f4fbf9;
  border-left: 4px solid #18bc9c;
  border-radius: 0 6px 6px 0;
  padding: 12px 18px;
  margin: 16px 0 20px;
  font-size: 0.88em;
  color: #444;
}

pre {
  background: #f8f9fa;
  border: 1px solid #e3e6ea;
  border-radius: 5px;
  font-size: 0.87em;
}
</style>

<a href="https://sung2021.github.io/" class="back-btn"><i class="fa fa-arrow-left"></i> Back to Main Page</a>  


# 1. Raw data processing  

**cellranger** : 10x Genomics's Chromium system to process single cell RNA-seq raw data. Easy to use but need large computing resources and Linux environment.   
Input : Fastq files, cellranger genome index  

## cellranger count/aggr  
```{bash}
# cellranger count for transcriptome data processing 
cellranger count --id=SampleA_GEX \
--transcriptome=/path/to/reference/transcriptomes/Mouse_GEX_2020/refdata-gex-mm10-2020-A \
--fastqs=/path/to/raw_data/SampleA \
--sample=SampleA \
--expect-cells=10000

# cellranger aggr to aggregate multiple cellranger count output to one file 
cellranger aggr --id=Aggregate_GEX \
--csv=/path/to/aggregate_info/Aggregate_GEX.csv

```


## Aggregate_GEX.csv example  
```{bash}
library_id,molecule_h5
SampleA,/path/to/SampleA/outs/molecule_info.h5
SampleB,/path/to/SampleB/outs/molecule_info.h5
SampleC,/path/to/SampleC/outs/molecule_info.h5

```

## Output  

**Aggregated Feature-Barcode Matrices**   

matrix.mtx   
genes.tsv   
barcodes.tsv   




# 2. Matrices processing and filtering     

## Load packages   
```{r, echo=T}
library(dplyr)
library(ggplot2)
library(reshape)
library(dplyr)
```


# Sample data  

Data import from the cellranger output (above).  
Create Seurat Object  
```{r, }
library(Seurat)
obj.raw = Read10X('raw_data/Path/filtered_feature_bc_matrix/')
obj.srt = CreateSeuratObject(counts = obj.raw, project = 'project_name')
```


# Doublet removal (by scrublet)  

scrublet runs in python env.  
Input file is the count matrix from seurat object.  


```{python}
import numpy as np
import pandas as pd
import scanpy as sc
import scrublet as scr

def run_scrublet(input_file, output_file, expected_doublet_rate=0.1):
    df = pd.read_csv(input_file, header=0, index_col=0)
    adata = sc.AnnData(df)
    
    # Set the expected_doublet_rate parameter
    sc.external.pp.scrublet(adata, expected_doublet_rate=expected_doublet_rate)
    
    # Save the observation (results) to CSV
    adata.obs.to_csv(output_file)

input_file = 'input.count.csv'
output_file = 'output.scr.csv'

# Adjust the expected_doublet_rate parameter
expected_doublet_rate = 0.1
run_scrublet(input_file, output_file, expected_doublet_rate)
```
<br>

Add the scrublet output information to the seurat object 
```{r}
## import scrublet result
df= read.csv('path/output.scr.csv', row.names = 1)

# If the previous process changes the name of cell ID, 
# Use the following code to correct it. 
rownames(df) = gsub(pattern = '_', replacement = '-' ,rownames(df))

## add doublet info to the srt obj

obj.srt[['doublet_score']] = df[rownames(obj.srt@meta.data),]$doublet_score
obj.srt[['predicted_doublet']] =df[rownames(obj.srt@meta.data),]$predicted_doublet

obj.srt %>% saveRDS("path/saved.obj.rds")

# You might want to check the number of doublets in the data
obj.srt@meta.data %>% select(predicted_doublet) %>% table()
```


# 3. Data Analysis   

# Perform default analysis 
**This process includes the following.**  

FindVariableFeatures("vst" method)   
NormalizeData   
ScaleData   
RunPCA   
FindNeighbors   
FindClusters   
RunUMAP   


```{r}
## perform default analysis
perform_default_analysis <- function(obj.srt, n_features = 2000, n_pcs = 20, 
                                     dims_for_neighbors = 1:20, 
                                     resolutions = c(0.1,0.5,1), 
                                     umap_dims = 1:20) {
  # Step 1: Find variable features
  obj.srt <- FindVariableFeatures(obj.srt, 
                                  selection.method = 'vst', 
                                  nfeatures = n_features)
  
  # Step 2: Scale and normalize data
  all_genes <- rownames(obj.srt)
  obj.srt <- NormalizeData(obj.srt)
  obj.srt <- ScaleData(obj.srt, features = all_genes)
  
  # Step 3: Run PCA
  obj.srt <- RunPCA(obj.srt, 
                    features = VariableFeatures(object = obj.srt), npcs = n_pcs)
  
  # Step 4: Find neighbors
  obj.srt <- FindNeighbors(obj.srt, dims = dims_for_neighbors)
  
  # Step 5: Find clusters
  obj.srt <- FindClusters(obj.srt, resolution = resolutions)
  
  # Step 6: Run UMAP
  obj.srt <- RunUMAP(obj.srt, dims = umap_dims)
  
  # Return the Seurat object with analysis results
  return(obj.srt)
}

# apply
obj.srt <- perform_default_analysis(obj.srt)
```



# Integration (Seurat v5 + Harmony)  

Seurat v5 uses `IntegrateLayers()` for batch correction. Harmony is recommended for most datasets due to its speed and robustness.  
Reference: https://satijalab.org/seurat/articles/seurat5_integration  
<br>

```{r}
library(harmony)

# Define batch variable
batch_var <- "Timepoint"

# Step 1: Split RNA assay into layers by batch (Seurat v5)
obj.srt[["RNA"]] <- split(obj.srt[["RNA"]], f = obj.srt@meta.data[[batch_var]])

# Step 2: Standard preprocessing
obj.srt <- NormalizeData(obj.srt)
obj.srt <- FindVariableFeatures(obj.srt)
obj.srt <- ScaleData(obj.srt)
obj.srt <- RunPCA(obj.srt, npcs = 30)

# Step 3: Harmony integration via IntegrateLayers
obj.srt <- IntegrateLayers(
  object = obj.srt,
  method = HarmonyIntegration,
  orig.reduction = "pca",
  new.reduction = "harmony",
  group.by.vars = batch_var,
  verbose = FALSE
)

# Step 4: Re-join layers after integration
obj.srt <- JoinLayers(obj.srt)

# Step 5: Downstream analysis using harmony reduction
obj.srt <- FindNeighbors(obj.srt, reduction = "harmony", dims = 1:30)
obj.srt <- FindClusters(obj.srt, resolution = c(0.1, 0.2, 0.5, 0.8))
obj.srt <- RunUMAP(obj.srt, reduction = "harmony", dims = 1:30)

obj.srt %>% saveRDS("path/saved.obj.rds")

# Visualization
DimPlot(obj.srt, reduction = "umap", group.by = batch_var)
```

**Other integration methods**  

RPCA (Seurat built-in, faster than CCA): use `method = RPCAIntegration`  
LIGER: https://github.com/welch-lab/liger  
<br>

# Cell cycle scoring 

```{r}
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
obj.srt <- CellCycleScoring(obj.srt, 
                            s.features = s.genes, g2m.features = g2m.genes)
```


# Cell type annotation by scType  

URL: https://github.com/IanevskiAleksandr/sc-type  


```{r}
# load gene set preparation function
source("https://raw.githubusercontent.com/IanevskiAleksandr/sc-type/master/R/gene_sets_prepare.R")
# load cell type annotation function
source("https://raw.githubusercontent.com/IanevskiAleksandr/sc-type/master/R/sctype_score_.R")

# DB file
db_ <- "https://raw.githubusercontent.com/IanevskiAleksandr/sc-type/master/ScTypeDB_full.xlsx";
tissue <- "Immune system" # e.g. Immune system,Pancreas,Liver,Kidney,Brain,Lung

# prepare gene sets
gs_list <- gene_sets_prepare(db_, tissue)

# extract scaled scRNA-seq matrix
scRNAseqData_scaled <- as.matrix(obj.srt[["RNA"]]@scale.data)

# run ScType
es.max <- sctype_score(scRNAseqData = scRNAseqData_scaled, 
                       scaled = TRUE, 
                       gs = gs_list$gs_positive, 
                       gs2 = gs_list$gs_negative)

# Merge by cluster
cL_results <- do.call("rbind", 
  lapply(unique(obj.srt@meta.data$seurat_clusters), function(cl) {
    cl_data <- obj.srt@meta.data[obj.srt@meta.data$seurat_clusters == cl, ]
    es_sum <- rowSums(es.max[, rownames(cl_data)])
    es_sorted <- sort(es_sum, decreasing = TRUE)
    top_es <- head(data.frame(cluster = cl, 
                              type = names(es_sorted), 
                              scores = es_sorted, 
                              ncells = nrow(cl_data)), 10)
    return(top_es)
  })
)

# Calculate top scores by cluster
sctype_scores <- cL_results %>% 
  group_by(cluster) %>% 
  top_n(n = 1, wt = scores)

# Set low-confidence clusters to "Unknown"
threshold <- sctype_scores$ncells / 4
sctype_scores$type[sctype_scores$scores < threshold] <- "Unknown"

```


```{r}
# overlay the identified cell types on UMAP plot
obj.srt@meta.data$sctype_classification = ""

# Assign cell type classification to meta data
unique_clusters <- unique(sctype_scores$cluster)

for (j in unique_clusters) {
  # Extract the type for the current cluster
  cl_type <- sctype_scores[sctype_scores$cluster == j, ]
  type_label <- as.character(cl_type$type[1])
  
  # Assign type label to all cells in the cluster
  obj.srt@meta.data$sctype_classification[obj.srt@meta.data$seurat_clusters == j] <- type_label
}

# Draw plot 
DimPlot(obj.srt, reduction = "umap", label = TRUE, repel = TRUE, 
        group.by = 'sctype_classification')        


```


# AUC (Area Under the Curve)   

To assess the activation of gene sets (custom) in assigned cell types (by scType), I calculate AUC scores by determining the level of activation of specific gene sets in each cell.  


```{r}
library(Seurat)
library(AUCell)
library(GSEABase)
```

```{r}
# import data from single-cell experiment
# Seurat v5: use LayerData() instead of @assays$RNA@data
exprMatrix <- LayerData(obj.srt, assay = "RNA", layer = "data")
# Convert to sparse:
exprMatrix <- as(exprMatrix, "dgCMatrix")
```

```{r}
# Create genesets to test
# getting them from msigDB or make custom ones 
gmt 
```

## Build gene-expression rankings for each cell   
```{r}
cells_rankings <- AUCell_buildRankings(exprMatrix, plotStats=TRUE)
```

## Calculate enrichment for the gene signatures (AUC)   
```{r}
cells_AUC <- AUCell_calcAUC(gmt, cells_rankings)
```

## Determine the cells with the given gene signatures or active gene sets   
```{r}
cells_assignment <- AUCell_exploreThresholds(cells_AUC, plotHist=TRUE, assign=TRUE) 
```

## Threshold check  
```{r}
# check the length inside 
n <- length(cells_assignment)

aucThr <- matrix(nrow = n, ncol = 1) # 필요한 경우 열의 수를 조정하세요.

# This is an example rownames
rownames(aucThr) <- c("B", "Fibroblast", "Myeloid", "T_NK", "Tumor")[1:n]

for (i in 1:n) {
  aucThr[i, ] <- cells_assignment[[i]]$aucThr$selected
}

# Data frame 
aucThr %>% data.frame() %>% DT::datatable()
```

## Add AUC output to metadata   
```{r}
# Add the calculated AUC scores as metadata to the Seurat object
obj.srt <- AddMetaData(obj.srt, data.frame(t(cells_AUC@assays@data$AUC)))

# This is an additional part
Category = c("B" ,"Fibroblast","Myeloid","T_NK","Tumor")

for (cat in Category) {
  rows <- cells_assignment[[cat]]$assignment
  auc_column_name <- paste0("AUC_designated_", cat)
  obj.srt@meta.data[, auc_column_name] <- "FALSE"
  obj.srt@meta.data[rows, auc_column_name] <- "TRUE"
}
```

## Visualize AUC score in UMAP   

**Note**   

Once obtaining AUC output, I check the information that can be transferred to seurat object metadata. There should be various visualization methods for the transferred data. I usually use UMAP visualization to check the general patterns of them.      
```{r}
DimPlot(obj.srt, group.by = "AUC_designated_my_category")
FeaturePlot(obj.srt, features = "AUC_score_for_my_geneset")
```


# Marker gene analysis  

## Find markers  
```{r}
# Find markers based on the selected cluster 
Idents(obj.srt) = 'RNA_snn_res.0.1'
fc=1.2
all.markers = FindAllMarkers(obj.srt, logfc.threshold = log2(fc), 
                             only.pos = T, test.use = 'wilcox',min.pct = 0.5)
```

## Create downloadable link of markers 
```{r}
all.markers %>% DT::datatable(extensions = "Buttons", 
                options = list(autoWidth = TRUE, 
                               fixedHeader = TRUE,
                               dom="Bfrtip", buttons=c("csv","excel")))
```

## Visualize the marker gene expression {.tabset} 

### Dotplot
```{r}
gs = genes_selected # Selected genes from the marker list

obj.srt$rev.res.0.1 = factor(obj.srt$RNA_snn_res.0.1, 
                             levels = rev(levels(obj.srt$RNA_snn_res.0.1)))

DotPlot(obj.srt, group.by = "rev.res.0.1", 
        features = gs, cols = c("grey","red")) +RotatedAxis()
```

### Heatmap  
```{r}
gs = genes_selected # Selected genes from the marker list
DoHeatmap(obj.srt, features = gs, group.by = "RNA_snn_res.0.1")
```


# DEG analysis between two conditions  
## Find DEGs   
```{r}
# Generating DEGs bewteen condition 1 and condition 2
DEG.two_groups= function(obj.srt,cluster=0, cond1="ADU", cond2="CTL"){
  obj.srt@meta.data=obj.srt@meta.data %>% 
    mutate(compare = ifelse(RNA_snn_res.0.1 == cluster & sample == cond1, 
                            paste0(cluster,"_", cond1), 
                            ifelse(RNA_snn_res.0.1 == cluster & sample == cond2, 
                                   paste0(cluster,"_", cond2), "others")))
  g2=paste0(cluster,"_", cond1)
  g1=paste0(cluster,"_", cond2)
  Idents(obj.srt) = 'compare'
  logfc=log2(1)
  mks =FindMarkers(obj.srt, ident.1 = g2, ident.2 = g1, 
                   logfc.threshold = logfc, assay = "count")
  pval=0.05
  fc=1.2
  mks = mks %>% 
    mutate(DE=ifelse(avg_log2FC >= log2(fc) & p_val_adj < pval, 'UP',
                                 ifelse(avg_log2FC <= -log2(fc) & p_val_adj < pval, 'DN',
                                        'no_sig')))
  mks$DE = factor(mks$DE, levels = c('UP','DN','no_sig'))
  
  mks$gene = rownames(mks)
  mks =mks %>% mutate(labels= ifelse(DE == 'UP', gene, ifelse(DE=='DN',gene,'')))
  mks =mks %>% arrange(desc(avg_log2FC))
  return(mks)
}

```

```{r}
# Application  
deg.cluster0 = DEG.two_groups(obj.srt = obj.srt, cluster = 0)
deg.cluster1 = DEG.two_groups(obj.srt = obj.srt, cluster = 1)
deg.cluster2 = DEG.two_groups(obj.srt = obj.srt, cluster = 2)
```

## Visualize the DEG output 
```{r}
# Volcano plot   
input.data = deg.cluster0
# Draw plot 
input.data %>% 
  ggplot(aes(avg_log2FC, -log10(p_val_adj), color=DE)) + 
  geom_point(size=1, alpha=0.5) + 
  scale_color_manual(values = c('red','blue','grey')) +
  theme_classic() +
  geom_vline(xintercept = c(-log2(fc),log2(fc)), color='grey') +
  geom_hline(yintercept = -log10(0.05),color='grey') +
  guides(colour = guide_legend(override.aes = list(size=5))) +
  ggtitle(paste0("Plot Title here")) +
  geom_text(aes(label = labels), size = 2.5, 
            show.legend = FALSE, hjust = 0, nudge_x = 0.01) +
  ggeasy::easy_center_title() 

```


# Gene set enrichment analysis of DEGs  

This analysis uses hallmark pathways for the analaysis.  
```{r}
library(clusterProfiler)
hallmark <- msigdbr::msigdbr(species = "Mus musculus", category = "H") %>% 
  dplyr::select(gs_name, gene_symbol)

perform_GSEA <- function(res, ref, pvalueCutoff = 1) {
  ranking <- function(res) {
    df <- res$avg_log2FC
    names(df) <- rownames(res)
    df <- sort(df, decreasing = TRUE)
    return(df)
  }
  
  ranked.res <- ranking(res)
  
  x <- clusterProfiler::GSEA(geneList = ranked.res,
                             TERM2GENE = ref,
                             pvalueCutoff = pvalueCutoff,
                             pAdjustMethod = "BH",
                             verbose = TRUE,
                             seed = TRUE)
  
  result <- x@result %>% arrange(desc(NES))
  result <- result[, c('NES', 'pvalue', 'p.adjust', 'core_enrichment', 'ID')]
  return(result)
}

# Application 
gsea.res = perform_GSEA(res = deg.cluster3, ref = hallmark)

```

## GSEA output visualization {.tabset}

### NES plot  
```{r}
# GESA Plot 
gsea_nes_plot =function(gsea.res, title){
  gsea.res %>% ggplot(aes(reorder(ID, NES), NES)) +
    geom_col(aes(fill=p.adjust)) +
    coord_flip() +
    labs(x="Pathway", y="Normalized Enrichment Score",
         title= "GSEA") + 
    theme_classic() +
    scale_fill_gradient(low = 'red', high = '#E5E7E9') +
    theme(axis.text.x= element_text(size=5, face = 'bold'),
          axis.text.y= element_text(size=6, face = 'bold'), 
          axis.title =element_text(size=10)) +ggtitle(title)
}

# Application 
gsea_nes_plot(gsea.res[gsea.res$p.adjust <= 0.05,], title = "Cluster3")
```


### Enrichment plot for the individual pathway  

```{r}
id1 = "GOBP_T_CELL_ACTIVATION" # Selected pathway" 
enrichplot::gseaplot2(gsea.res, geneSetID = id1, title = id1)

```


# Trajectory Analysis (Monocle3)  

Trajectory analysis infers developmental or activation paths across cells. Monocle3 works natively with Seurat objects via `SeuratWrappers`.  
Reference: https://cole-trapnell-lab.github.io/monocle3/  
<br>

```{r}
library(monocle3)
library(SeuratWrappers)

# Convert Seurat object to CellDataSet
cds <- as.cell_data_set(obj.srt)

# Pre-process (skip if already done in Seurat)
cds <- preprocess_cds(cds, num_dim = 50)

# Transfer UMAP coordinates from Seurat
reducedDim(cds, type = "UMAP") <- obj.srt@reductions$umap@cell.embeddings[, 1:2]
cds@int_colData@listData$reducedDims$UMAP <- obj.srt@reductions$umap@cell.embeddings[, 1:2]

# Cluster and learn trajectory graph
cds <- cluster_cells(cds)
cds <- learn_graph(cds)

# Order cells along trajectory
# Use order_cells(cds) interactively, or specify root cells:
# cds <- order_cells(cds, root_cells = root_cell_ids)
cds <- order_cells(cds)

# Visualize pseudotime
plot_cells(cds,
           color_cells_by = "pseudotime",
           label_cell_groups = FALSE,
           label_leaves = FALSE,
           label_branch_points = FALSE,
           graph_label_size = 1.5)
```


# Imputation  

MAGIC (Cell, 2018)   
https://github.com/KrishnaswamyLab/MAGIC   
Markov Affinity-based Graph Imputation of Cells (MAGIC) is an algorithm for denoising high-dimensional data most commonly applied to single-cell RNA sequencing data. MAGIC learns the manifold data, using the resultant graph to smooth the features and restore the structure of the data.   
<br>

```{python}
# Import the necessary libraries.
import pandas as pd
import magic
import csv

# Load the CSV file into a DataFrame.
data_path = '~/path/data.csv'
x = pd.read_csv(data_path)

# Remove the 'Unnamed: 0'
del x['Unnamed: 0']

# Initialize the MAGIC algorithm. the number of nearest neighbors (knn) is set to 5, and the diffusion time (t) is set to 3.
magic_operator = magic.MAGIC(knn=5, t=3)

# Apply the MAGIC algorithm to the data.
x_magic = magic_operator.fit_transform(x)

# Save the data to a CSV file.
output_path = '~/path/data.magic.csv'
x_magic.to_csv(output_path)

```



<!-- # Further analysis -->



<!-- <br><br> -->
<!-- More example analyses will be updated.   -->

<!-- <br><br> -->

